Pre-/Post-Trikafta in CFRD

### 2022-10-07 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

```{r include=FALSE}
# Put everything together
chco <- chco %>% select(all_of(t1_variables))
mon <- mon %>% select(all_of(t1_variables))
um <- um %>% select(all_of(t1_variables))
uw <- uw %>% select(all_of(t1_variables))
df <- bind_rows(chco, mon, um, uw) %>% arrange(`CFF ID`, Time)
# Age group
df$`Age Group` <- cut(df$Age, breaks = c(-Inf, 20, Inf), 
                      labels = c("Pedatric", "Adult"), right = F)
# Race and sex
df$Race <- factor(df$Race,
  levels = c(
    "Black or African American", "Caucasian",
    "Other", "Some other race", "Two or more races", "White"
  ),
  labels = c(
    "Black", "Caucasian",
    "Other", "Other", "Other", "Caucasian"
  )
)
df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
# FEV calculations
age <- df$Age
height <- df$Height / 100
race <- df$Race
old <- c("Caucasian", "Black", "Other")
new <- c(1, 2, 5)
race <- new[match(race, old)]
gender <- as.numeric(df$Sex)
df$ppFEV1 <- pctpred_GLI(age, height, gender = gender, ethnicity = race, FEV1 = df$FEV1)
df$ppFVC <- pctpred_GLI(age, height, gender = gender, ethnicity = race, FVC = df$FVC)
# Genotype severity
df$Genotypes1 <- get_cf_genotype_class_severity(df$Genotypes1)
df$Genotypes2 <- get_cf_genotype_class_severity(df$Genotypes2)
df <- df %>% unite(Genotype, Genotypes1, Genotypes2)
df$Genotype <- factor(df$Genotype,
  levels = c(
    "Mild_Severe", "Severe_Mild", "Severe_Severe",
    "Severe_Unknown", "Unknown_Unknown"
  ),
  labels = c("Low", "Low", "High", "Unknown", "Unknown")
)
# Other little cleaning things
df <- df %>% filter(Age > 0, !is.na(Time))
df$`G-tube in past 12 months`[is.na(df$`G-tube in past 12 months`)] <- "No"
# Pre-/post-modulator
df$`Pre or Post` <- factor(sign(df$Time), levels = -1:1, labels = c("Pre", "Pre", "Post"))
# Add insulin information
chco_ins <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
chco_ins$`CFF ID` <- chco_a1c$`CFF ID`[match(chco_ins$MRN, chco_a1c$MRN)]
chco_ins <- chco_ins %>%
  select(`CFF ID`, `Pre or Post`, `Total Daily Dose of insulin (units)`) %>%
  rename(TDD = `Total Daily Dose of insulin (units)`) %>%
  filter(!is.na(`CFF ID`))
mon_ins <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_insulin.csv", show_col_types = F)
mon_ins <- mon_ins %>%
  select(`CFF ID`, `Pre or Post`, `Total Daily Dose of insulin (units)`) %>%
  rename(TDD = `Total Daily Dose of insulin (units)`)
uw_ins <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_insulin.csv", show_col_types = F)
uw_ins <- uw_ins %>%
  select(`CFF ID`, `Pre or Post`, `Total Daily Dose of insulin (units)`) %>%
  rename(TDD = `Total Daily Dose of insulin (units)`)
insulin <- bind_rows(chco_ins, mon_ins, uw_ins)
# Join
df <- left_join(df, insulin)
df$`Pre or Post` <- factor(df$`Pre or Post`, levels = c("Pre", "Post"))
# Get pre value closest to start, post value closest to 1 year from start
df$select_time = df$Time
df$select_time[df$`Pre or Post`=="Post"] = 365 - df$select_time[df$`Pre or Post`=="Post"]
df <- df %>%
  arrange(`CFF ID`,`Pre or Post`, select_time) %>%
  group_by(`CFF ID`, `Pre or Post`) %>%
  summarise(
    `Num. A1c Values` = sum(!is.na(HbA1c)),
    across(everything(), ~ last(na.omit(.x))),
  .groups = "keep") %>%
  filter(!is.na(HbA1c))
keep = df %>% group_by(`CFF ID`) %>% summarise(n=n()) %>% filter(n==2) %>% pull(`CFF ID`)
df = df %>% filter(`CFF ID` %in% keep)
```

```{r include=FALSE}
cgm_vars <- c(
  "average_sensor", "min_sensor", "max_sensor", "percent_time_over_120",
  "percent_time_over_140", "percent_time_over_200", "excursions_over_140",
  "excursions_over_200", "percent_time_under_60", "percent_time_under_70",
  "average_auc_per_day", "standard_deviation", "r_mage"
)
# Add CGM information
# CHCO
chco_cgm <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_cgm.csv", show_col_types = F)
chco_cgm$`Pre or Post` <- factor(sapply(str_split(chco_cgm$subject_id, "_"), "[", 2))
levels(chco_cgm$`Pre or Post`) <- c("Post", "Pre", "Pre")
chco_cgm$`CFF ID` <- sapply(str_split(chco_cgm$subject_id, "_"), "[", 1)
chco_cgm <- chco_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(cgm_vars))
# Montana
mon_cgm <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_cgm.csv", show_col_types = F)
mon_cgm$`Pre or Post` <- factor(sapply(str_split(mon_cgm$subject_id, "_"), "[", 2))
levels(mon_cgm$`Pre or Post`) <- c("Post", "Post", "Pre", "Pre")
mon_cgm$`CFF ID` <- sapply(str_split(mon_cgm$subject_id, "_"), "[", 1)
mon_cgm <- mon_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(cgm_vars))
# UW
uw_cgm <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cgm.csv", show_col_types = F)
uw_cgm$`Pre or Post` <- factor(sapply(str_split(uw_cgm$subject_id, "_"), "[", 2))
levels(uw_cgm$`Pre or Post`) <- c("Post", "Pre")
uw_cgm$`CFF ID` <- sapply(str_split(uw_cgm$subject_id, "_"), "[", 1)
uw_cgm <- uw_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(cgm_vars))
# Put together
cgm <- bind_rows(chco_cgm, mon_cgm, uw_cgm)
# Add to DF
cgm$`CFF ID` = as.numeric(cgm$`CFF ID`)
df = left_join(df,cgm)
```

>>>>

```{r include=FALSE}
# Put everything together
chco <- chco %>% select(all_of(t1_variables))
mon <- mon %>% select(all_of(t1_variables))
um <- um %>% select(all_of(t1_variables))
uw <- uw %>% select(all_of(t1_variables))
df <- bind_rows(chco, mon, um, uw) %>% arrange(`CFF ID`, Time)
# Age group
df$`Age Group` <- cut(df$Age,
  breaks = c(-Inf, 20, Inf),
  labels = c("Pedatric", "Adult"), right = F
)
# Race and sex
df$Race <- factor(df$Race,
  levels = c(
    "Black or African American", "Caucasian",
    "Other", "Some other race", "Two or more races", "White"
  ),
  labels = c(
    "Black", "Caucasian",
    "Other", "Other", "Other", "Caucasian"
  )
)
df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
# FEV calculations
age <- df$Age
height <- df$Height / 100
race <- df$Race
old <- c("Caucasian", "Black", "Other")
new <- c(1, 2, 5)
race <- new[match(race, old)]
gender <- as.numeric(df$Sex)
df$ppFEV1 <- pctpred_GLI(age, height, gender = gender, ethnicity = race, FEV1 = df$FEV1)
df$ppFVC <- pctpred_GLI(age, height, gender = gender, ethnicity = race, FVC = df$FVC)
# Genotype severity
df$Genotypes1 <- get_cf_genotype_class_severity(df$Genotypes1)
df$Genotypes2 <- get_cf_genotype_class_severity(df$Genotypes2)
df <- df %>% unite(Genotype, Genotypes1, Genotypes2)
df$Genotype <- factor(df$Genotype,
  levels = c(
    "Mild_Severe", "Severe_Mild", "Severe_Severe",
    "Severe_Unknown", "Unknown_Unknown"
  ),
  labels = c("Low", "Low", "High", "Unknown", "Unknown")
)
# Other little cleaning things
df <- df %>% filter(Age > 0, !is.na(Time))
df$`G-tube in past 12 months`[is.na(df$`G-tube in past 12 months`)] <- "No"
# Pre-/post-modulator
df$`Pre or Post` <- factor(sign(df$Time), levels = -1:1, labels = c("Pre", "Pre", "Post"))
# Add insulin information
chco_ins <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
chco_ins$`CFF ID` <- chco_a1c$`CFF ID`[match(chco_ins$MRN, chco_a1c$MRN)]
chco_ins <- chco_ins %>%
  select(`CFF ID`, `Pre or Post`, `Total Daily Dose of insulin (units)`) %>%
  rename(TDD = `Total Daily Dose of insulin (units)`) %>%
  filter(!is.na(`CFF ID`))
mon_ins <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_insulin.csv", show_col_types = F)
mon_ins <- mon_ins %>%
  select(`CFF ID`, `Pre or Post`, `Total Daily Dose of insulin (units)`) %>%
  rename(TDD = `Total Daily Dose of insulin (units)`)
uw_ins <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_insulin.csv", show_col_types = F)
uw_ins <- uw_ins %>%
  select(`CFF ID`, `Pre or Post`, `Total Daily Dose of insulin (units)`) %>%
  rename(TDD = `Total Daily Dose of insulin (units)`)
insulin <- bind_rows(chco_ins, mon_ins, uw_ins)
# Join
df <- left_join(df, insulin)
df$`Pre or Post` <- factor(df$`Pre or Post`, levels = c("Pre", "Post"))
# Get pre value closest to start, post value closest to 1 year from start
df$select_time <- df$Time
df$select_time[df$`Pre or Post` == "Post"] <- 365 - df$select_time[df$`Pre or Post` == "Post"]
df <- df %>%
  arrange(`CFF ID`, `Pre or Post`, select_time) %>%
  group_by(`CFF ID`, `Pre or Post`) %>%
  summarise(
    `Num. A1c Values` = sum(!is.na(HbA1c)),
    across(everything(), ~ last(na.omit(.x))),
    .groups = "keep"
  ) %>%
  filter(!is.na(HbA1c))
keep <- df %>%
  group_by(`CFF ID`) %>%
  summarise(n = n()) %>%
  filter(n == 2) %>%
  pull(`CFF ID`)
df <- df %>% filter(`CFF ID` %in% keep)
```

```{r include=FALSE}
cgm_vars <- c(
  "average_sensor", "min_sensor", "max_sensor", "percent_time_over_120",
  "percent_time_over_140", "percent_time_over_200", "excursions_over_140",
  "excursions_over_200", "percent_time_under_60", "percent_time_under_70",
  "average_auc_per_day", "standard_deviation", "r_mage"
)
# Add CGM information
# CHCO
chco_cgm <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_cgm.csv", show_col_types = F)
chco_cgm$`Pre or Post` <- factor(sapply(str_split(chco_cgm$subject_id, "_"), "[", 2))
levels(chco_cgm$`Pre or Post`) <- c("Post", "Pre", "Pre")
chco_cgm$`CFF ID` <- sapply(str_split(chco_cgm$subject_id, "_"), "[", 1)
chco_cgm <- chco_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(cgm_vars))
# Montana
mon_cgm <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_cgm.csv", show_col_types = F)
mon_cgm$`Pre or Post` <- factor(sapply(str_split(mon_cgm$subject_id, "_"), "[", 2))
levels(mon_cgm$`Pre or Post`) <- c("Post", "Post", "Pre", "Pre")
mon_cgm$`CFF ID` <- sapply(str_split(mon_cgm$subject_id, "_"), "[", 1)
mon_cgm <- mon_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(cgm_vars))
# UW
uw_cgm <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cgm.csv", show_col_types = F)
uw_cgm$`Pre or Post` <- factor(sapply(str_split(uw_cgm$subject_id, "_"), "[", 2))
levels(uw_cgm$`Pre or Post`) <- c("Post", "Pre")
uw_cgm$`CFF ID` <- sapply(str_split(uw_cgm$subject_id, "_"), "[", 1)
uw_cgm <- uw_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(cgm_vars))
# Put together
cgm <- bind_rows(chco_cgm, mon_cgm, uw_cgm)
# Add to DF
cgm$`CFF ID` <- as.numeric(cgm$`CFF ID`)
df <- left_join(df, cgm)
```
